Fréchet Distance

### 2022-11-30 docs/paper.tex
finish text of original algorithm

% TODO: Why allow it? Better algorithms?
\subsection{Weak Fréchet distance}
The \textit{weak Fréchet distance} slightly modifies the continuous version by allowing 
going backwards.

@@

We observe that there exists 
a monotonically increasing curve from $(0,0)$ to $(p,q)$ in $F_\varepsilon$
if and only if the Fréchet distance between $P$ and $Q$ is less than or equal to $\varepsilon$,
i.e., $\delta_F(P,Q) \leq \varepsilon$.
Given such a curve, the walking speed functions $\alpha$ and $\beta$ from \autoref{def_frechet_distance}
can be easily obtained by rescaling the parameter ranges.

Computing the complete free space is very compute intensive 
because each point in the continuous space needs to be computed. 
Using the convexity property of each cell in the free space,
we can make another observation requiring only a few points to be computed per cell. 
These points define the legal regions on each boundary side of the cell, see \autoref{fig_legal_boundary_intervals_of_a_cell}. 
We don't need to know the exact shape of the free space 
because we travel in a monotonically increasing way through a convex space,
hence never entering any illegal regions.

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption[Legal intervals on free space cell boundary]{Legal intervals on the boundaries of a free space cell.}
  \label{fig_legal_boundary_intervals_of_a_cell}
\end{figure}

>>>>

% TODO: Why allow it? Better algorithms?
\subsection{Weak Fréchet distance}
The \textit{weak Fréchet distance} (non-monotone Fréchet distance) slightly modifies the continuous version by allowing 
going backwards.

@@

Computing the complete free space $F_\varepsilon$ is very compute intensive 
because each point in the continuous space needs to be computed. 
Using the convexity property of each cell in the free space,
we can observe that only a small constant number of points need to be computed per cell. 
These points define the legal regions on each boundary side of the cell, see \autoref{fig_legal_boundary_intervals_of_a_cell}. 
We don't need to know the exact shape of the free space 
because we travel in a monotonically increasing way through a convex space,
hence never entering any illegal regions.

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption[Legal intervals on free space cell boundary]{
    Legal intervals on the boundaries of a free space cell. 
    There are at most 8 intersection points for an ellipse and a rectangle.}
  \label{fig_legal_boundary_intervals_of_a_cell}
\end{figure}

Due to the constant number of intersection points required per cell, 
the free space of a single cell can be computed in $\mathcal{O}(1)$ time.
Therefore, the complete free space of all cells can be computed in $\mathcal{O}(pq)$
because there are $p-1$ (respectively $q-1$) edges in $P$ (respectively $Q$).

We observe that there exists 
a monotonically increasing curve from $(0,0)$ to $(p,q)$ in $F_\varepsilon$
if and only if the Fréchet distance between $P$ and $Q$ is less than or equal to $\varepsilon$,
i.e., $\delta_F(P,Q) \leq \varepsilon$.
Given such a curve, the walking speed functions $\alpha$ and $\beta$ from \autoref{def_frechet_distance}
can be easily obtained by rescaling the parameter ranges. 
Computing such a monotonically increasing curve can be done using dynamic programming 
by computing the bottom left most reachable points in each cell starting at $(0,0)$. 
In each step the solution for a single cell can be computed in constant time based on the 
solutions to previously computed cells. 
Therefore, the runtime is also $\mathcal{O}(pq)$. 
See \autoref{fig_free_space_curve} for an example curve. \cite{alt_computing_1995}

\begin{figure}[ht]
  \centering
  % TODO: Replace graphic with appropiate example
  \includegraphics[width=\textwidth]{images/hausdorff/hausdorff-distance-example.pdf}
  \caption{The monotonically increasing curve through the free space is computed by dynamic programming.}
  \label{fig_free_space_curve}
\end{figure}

% Actually computing the exact Fréchet distance
So far we discussed the decision problem whether $\delta_{F}(P,Q) \leq \varepsilon$
and observe that it is solvable in $\mathcal{O}(pq)$ time.
What we actually want to know is the exact value of the Fréchet distance.
A straight forward approach for this is to compute the value bit by bit 
using binary search in total $\mathcal{O}(pq \log (\text{``accuracy bits''}))$ time.
In practice this approach works really well 
and is often favored over other more complicated approaches. \cite{alt_computing_1995}

Another approach is to check all values of $\varepsilon$ where something ``interesting'' happens
until the smallest value of $\varepsilon$ is obtained where a monotonically increasing path 
from $(0,0)$ to $(p,q)$ in $F_\varepsilon$ exists.
We will call these the \textit{critical values} of $\varepsilon$, which are:
\begin{enumerate}
  \item the minimal value of $\varepsilon$ where $(0,0) \in F_\varepsilon$ and $(p,q) \in F_\varepsilon$.
  \item values of $\varepsilon$ where a new passage on the border edge 
        between two neighboring cells opens up.
  \item values of $\varepsilon$ where a new passage horizontal (respectively vertical) passage 
        between two non-neighboring cells opens up.
\end{enumerate}

There is one value of case 1, 
specifically the maximum of starting point to starting point and endpoint to endpoint distance,
$\mathcal{O}(pq)$ values of case 2, 
namely the number of vertex to edge distances,
and $\mathcal{O}(p^2q + pq^2)$ values of case 3,
particularly the number of times two vertices of one curve have the same distance to one edge of the other curve.
Determining all these values can be done in $\mathcal{O}(p^2q + pq^2)$ time 
because each value in each case can be computed in constant time, 
making the values of the third case dominate the time complexity.
% TODO: why does this simplyfication work
These values need to be sorted in $\mathcal{O}((p^2q + pq^2) \log (p^2q + pq^2)) = \mathcal{O}((p^2q + pq^2) \log (pq))$ time
in order to perform binary search in $\mathcal{O}(pq \log (p^2q + pq^2))$ time on them
using the algorithm for the decision problem.
The overall time required for this approach is dominated by the sorting step, 
thus $\mathcal{O}((p^2q + pq^2) \log (pq))$.
This time complexity can be decreased to $\mathcal{O}(pq \log (pq))$ 
using parametric search \cite{megiddo_applying_1983, cole_slowing_1987} to work faster in theory.
However, due to the specific internally used sorting network 
it is hard to implement this efficiently in practice. \cite{alt_computing_1995}
